---
title: "Statistics Homework"
output:
  html_document:
  toc: yes
toc_float: yes
link-citations: yes
bibliography: bibliography.bib
date: '2022-06-12'
---


  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(tidyverse)
library(gt)
library(ggplot2)
library(ggcorrplot)
library(cowplot)
library(GGally)
library(corrplot)
library(psych)

```



### INTRODUCTION



#### Here we have multiple questions: 
* How does the data look like and its visualization
* What variables correlate to each other 
* Quality research
* What other variables say about the quality
* Main differences between two families of wines




### How does data look like and what is its summary?

Head for white wine dataset


```{r data123, echo=FALSE}
white <- read.csv(file = 'C:/Users/Spelcina/Desktop/winequality-white.csv',sep=";");

white %>% head(5) %>% gt()

```


Head for red wine dataset


```{r data, echo=FALSE}
red <- read.csv(file = 'C:/Users/Spelcina/Desktop/winequality-red.csv',sep=";");

red %>% head(5) %>% gt()

```


Summary for red wine

```{r data2212, echo=FALSE}

summary(red)

```


Summary for white wine

```{r data222, echo=FALSE}

summary(white)

```





### We want to see what variables coorelate to each other


Correlation of variables of red wines:

```{r colleges, echo=FALSE}

cor(white)



```

Correlation of white wines

```{r college1s, echo=FALSE}


cor(red)

```


Red wine correlation matrix graph for easier visualization:


```{r scorer1s, warning=FALSE,echo=FALSE}

ggcorrplot(cor(red))


```


White wine correlation matrix


```{r scorers, warning=FALSE,echo=FALSE}

ggcorrplot(cor(white))


```

Here I want to plot residual sugar vs density and alcohol vs density for both wines

Red sugar vs density


```{r coqr132321, echo=FALSE}

red %>% ggplot(aes(x=residual.sugar,y=density ,group=1)) + geom_point() 

```

White RS vs density


```{r coqr135321, echo=FALSE}

white %>% ggplot(aes(x=residual.sugar,y=density ,group=1)) + geom_point() 

```


White alcohol vs density


```{r coqr137621, echo=FALSE}

white %>% ggplot(aes(x=alcohol,y=density ,group=1)) + geom_point() 

```


And red density vs alcohol


```{r qqq9989898, echo=FALSE}

red %>% ggplot(aes(x=alcohol,y=density ,group=1)) + geom_point() 

```


### We are interested in the quality of the wine, so we want to visualize it

Histogram for red wine quality

```{r qqr, echo=FALSE}
hist(red$quality) 

```

Histogram for white wine quality

```{r qqr1, echo=FALSE}
hist(white$quality)

```


Boxplot for red wines quality

```{r cow21r, echo=FALSE}

boxplot(red$quality)

```

Boxplot for white wines quality

```{r cowr, echo=FALSE}

boxplot(white$quality)

```


Red quality variance:

```{r tru11,warning=FALSE,echo=FALSE}
var(red$quality)


```

White wine variance:

```{r tru111,warning=FALSE,echo=FALSE}
var(white$quality)


```


We want see how alcohol and density relate to the quality for red

```{r coqr, echo=FALSE}

red %>% ggplot(aes(x=alcohol,y= density,group=1)) + geom_point(aes(colour = quality,size=1)) 

```

This wasn't as conclusive as I wished it to be


Now we can see the same thing also for white wine

```{r coqr1, echo=FALSE}

white %>% ggplot(aes(x=alcohol,y= density,group=1)) + geom_point(aes(colour = quality,size=1)) 

```
Now we are comparing red wines residual sugar with volatile acidity in regard to quality

```{r coqr11, echo=FALSE}

red %>% ggplot(aes(x=residual.sugar,y=volatile.acidity ,group=1)) + geom_point(aes(colour = quality,size=1)) 

```

We can see the less acidic it is and less sugar it has the more quality it has for red wine


We do the same thing also for white wine

```{r coqr1321, echo=FALSE}

white %>% ggplot(aes(x=residual.sugar,y=volatile.acidity ,group=1)) + geom_point(aes(colour = quality,size=1)) 

```


Here we dont see much significance

### Mean comparisons between wines


We want to find the wine with less residual sugar between these two wine families

My hypothesis here is that they do not vary much in terms of residual sugar

```{r tru,warning=FALSE,echo=FALSE}

t.test(white$residual.sugar, red$residual.sugar,var.equal=TRUE)

```


Now we want to test for volatile acidity, hypothesis is that there is no significant difference

```{r tru123,warning=FALSE,echo=FALSE}

t.test(white$volatile.acidity, red$volatile.acidity,var.equal=TRUE)

```

From this test we determine that red is more acidic than the white wine which is again surprising

Next variable we want to test is the density, with the same assumption

```{r tru121233,warning=FALSE,echo=FALSE}

t.test(white$density, red$density,var.equal=TRUE)

```

We again get the same results, that the true difference is still not 0. But the difference in means is not so big


Last variable I want to test is the PH value
```{r tru13223,warning=FALSE,echo=FALSE}

t.test(white$pH, red$pH,var.equal=TRUE)

```

Which again is not the same, but means aren't so far apart

Which gave me idea to compare all variables means from both means to truly understand the difference between wines

```{r tru123323,warning=FALSE,echo=FALSE}

qq<-colMeans(white)

qq2<-colMeans(red)

qq-qq2
```

These are the differences of means of both wine families. I know the values aren't scaled, but from looking at the data, we can conclude that the biggest difference is the total sulfur dioxide. 

### Conclusion

By this I want to conclude the analysis, and want to say that there are bigger differences than I expected. This can give insight when buying wine. As for statistical tools used, mostly it was pearson correlation, t-test and simple graphs and mean comparisons.
